### 2016-12-12 paper/chapters/introduction.tex
more on introduction

\section{Introduction}
\label{sec:introduction}
Generative models are one of the most promising ways to perform unsupervised learning (citation needed).
More specifically, generative modelling assumes a hidden structure $z$ explaining the input data $x$.
This assumption is reasonable given the following examples:\\

\begin{itemize}
  \item \textbf{Images} often have multiple thousands of dimensions, but the underlying structure is often far simpler. For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions. There are a few more dimensions, stroke width, cursive, position, size. But still far lower than 784. We will take a look at this example later on in \ref{sec:vae}.
  \item \textbf{Speech} or more precisely the recording of it is high-dimensional, while the information it carries (the spoken words) can be represented in a low-dimensional space.
  \item \textbf{...}
\end{itemize}

After a short overview of different approaches and methods and an introductory example with the sigmoid belief net (SBN), we will take a closer look at two promising frameworks, the variational autoencoder (VAE) and the generative adversarial network (GAN)

>>>>

\section{Introduction}
\label{sec:introduction}

"What I cannot create, I do not understand." -- Richard Feynman \\\\\\

% on building reasonable representations
One of the key motivations to generative modelling is the assumption that having a interpretable representation of the real world is important to building intelligent agents.
Real world data is often high-dimensional and noise and transforming this data into an compact representation is therefore an important and also challenging task to perform.
Generative models are ideal for this task since it has to understand the underlying structure to generate similar data.
Building this structure is called \textbf{representation learning} and is key to many generative models and how they perform.\\

With this assumption in mind we can take a look at some examples of input data.
\begin{itemize}
  \item \textbf{Images} commonly have thousands of dimensions, but the underlying structure is often far simpler. For example are 28x28 images of handwritten digits 784-dimensional but assuming one-hot encoding the most important information - the digit - it can be explained using 10 dimensions. There are a few more dimensions, stroke width, cursive, position, size. But still far lower than 784. We will take a look at this example later on in \ref{sec:vae}.
  \item \textbf{Speech} or more precisely the recording of it is high-dimensional, while the information it carries (the spoken words) can be represented in a low-dimensional space.
  \item \textbf{...}
\end{itemize}

To formalize this setting, it is often beneficial to view the structure as a probability distribution $p$ on the input data $x$ we build the structure 
%( another word!)
from: $p(x)$.

The assumption that $p(x)$ is dependent on some underlying structure can now be expressed as having the conditional probability $p(x|z)$ where $z$ are the causal factors.
We call $z$ also the latent variables.\\

%Inverting $p(x|z)$ yields $p(z|x)$ which is one of the main tasks of some generative models, for example the VAE \ref{sec:vae}.

After a short overview of different approaches and methods and an introductory example with the sigmoid belief net (SBN), we will take a closer look at two recent promising frameworks, the variational autoencoder (VAE \ref{sec:vae}) and the generative adversarial network (GAN \ref{sec:gan}).

%Generative models are one of the most promising ways to perform unsupervised learning (citation needed).
%More specifically, generative modelling assumes a hidden structure $z$ explaining the input data $x$.
%This assumption is reasonable given the following examples:\\
